---
title: "Project_4"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_section: true
---

```{r, message = FALSE, warning=FALSE}
require(survival)
require(survminer)
require(tidyr)
require(ranger)
require(psych)
require(dplyr)
require(ggplot2)
require(ggfortify)
require(coin)
require(GGally)
require(survminer)
require(feasts)
theme_set(theme_bw())
```

# Import dataset

```{r}
data("ovarian")
```

# EDA

## Data structure

```{r echo=FALSE}
dim(ovarian)
str(ovarian)
```

    All of variables type is numeric
    But according to the meaning of the variables, some of them are not numeric:
    - futime:	survival or censoring time
    - fustat:	censoring status
    - age:	in years
    - resid.ds:	residual disease present (1=no,2=yes)
    - rx:	treatment group
    - ecog.ps:	ECOG performance status (1 is better, see reference)

## Сonvert the type of variables to the correct type

### Rename columns

```{r}
names(ovarian)[4] <- "resid_disease" 
names(ovarian)[5] <- "treat_group" 
names(ovarian)[6] <- "ecog_status" 
```

### Change type

```{r}
ovarian$resid_disease <- factor(ovarian$resid_disease, labels = c("No", "Yes"))
ovarian$treat_group <- as.factor(ovarian$treat_group)
ovarian$ecog_status <- as.factor(ovarian$ecog_status)
```

```{r echo=FALSE}
str(ovarian)
```


## Check number of NA

```{r}
colSums(is.na(ovarian))
table(rowSums(is.na(ovarian)))
```

    Dataset have no NA

## Data analysis 

```{r}
ggplot(ovarian, aes(x=treat_group, fill=treat_group)) +
  geom_bar(alpha=0.6)  + 
  ggtitle("Count of patients in each treatment group") + 
  xlab("Treatment group") + labs(fill = "Treatment group")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

```{r}
paste0("Number of observations in 1 treatment group: ", nrow(subset(ovarian, treat_group==1)))
paste0("Number of observations in 2 treatment group: ", nrow(subset(ovarian, treat_group==2)))
```

    In each treatment group the same number of observations

```{r}
ggplot(ovarian, aes(y=age, x=treat_group, fill=treat_group)) +
  geom_boxplot(alpha=0.6)  + 
  ggtitle("Box plot of age colored by treatment group") + 
  xlab("Treatment group") + ylab("Age") + labs(fill = "Treatment group")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

    In the first group of therapy there are wider range of ages, and in the second group mainly people aged ~ 55-60 years(with 1 outlier)
    
    
```{r}
ggplot(ovarian, aes(x=ecog_status, fill=ecog_status)) +
  geom_bar(alpha=0.6)  + 
  ggtitle("Count of patients in each group of ECOG performance status") + 
  xlab("ECOG performance status") + labs(fill = "ECOG performance status")+
  theme(plot.title = element_text(size=12, face="bold", hjust = 0.7))
```

```{r}
paste0("Number of observations in 1 ECOG group: ", nrow(subset(ovarian, ecog_status==1)))
paste0("Number of observations in 2 ECOG group: ", nrow(subset(ovarian, ecog_status==2)))
```

    Number of observations in '1' ECOG performance status more than in '2'

```{r}
ggplot(ovarian, aes(x=ecog_status, y=age, fill=ecog_status)) +
  geom_boxplot(alpha=0.6)  + 
  ggtitle("Box plot of age colored by ECOG performance status") + 
  xlab("ECOG performance status") + ylab("Age") + 
  labs(fill = "ECOG performance status") +
  theme(plot.title = element_text(size=12, face="bold", hjust = 0.7))
```
    
    In the first group of ECOG status there are wider range of ages, and in the second group mainly people aged ~ 60 years(and 3 outliers observed)
    
    
```{r}
ggplot(ovarian, aes(x=resid_disease, fill=resid_disease)) +
  geom_bar(alpha=0.6)  + 
  ggtitle("Count of patients in each group of residual disease present") + 
  xlab("Residual disease present") + labs(fill = "Residual disease present")+
  theme(plot.title = element_text(size=12, face="bold", hjust = 0.7))
```

```{r}
paste0("Number of patients who presents residual disease: ", nrow(subset(ovarian, resid_disease=="Yes")))
paste0("Number of patients without residual disease: ", nrow(subset(ovarian, resid_disease=="No")))
```

    Number of patients who presents residual disease more then patient without residual disease

```{r}
ggplot(ovarian, aes(x=resid_disease, y=age, fill=resid_disease)) +
  geom_boxplot(alpha=0.6)  + 
  ggtitle("Box plot of age colored by residual disease present") + 
  xlab("Residual disease present") + ylab("Age") + 
  labs(fill = "Residual disease present") +
  theme(plot.title = element_text(size=12, face="bold", hjust = 0.7))
```
      
      The age of people who presence of residual disease is higher than of people without residual disease
      
# Task 1. Kaplan Meier Survival Curve

```{r}
km <- with(ovarian, Surv(futime, fustat))
head(km,26)
```

```{r}
km_fit <- survfit(Surv(futime, fustat) ~ 1, data=ovarian)
summary(km_fit)

autoplot(km_fit) + 
  ggtitle("Kaplan Meier survival curve") + 
  xlab("Time (days)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```
    
    Human survival decreases over time, investigate what factors affect survival.
    
## Build a linear model that takes into account the change in survival based on a factors over time

### Residual disease present

```{r}
km_fit_resid_disease <- survfit(Surv(futime, fustat) ~ resid_disease , 
                                data=ovarian)
summary(km_fit_resid_disease)

autoplot(km_fit_resid_disease) + 
  ggtitle("Kaplan Meier survival curve based on residual disease present") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```
      
    The survival rate of people with residual disease is lower than without them
    
### Treatment group

```{r}
km_fit_trt <- survfit(Surv(futime, fustat) ~ treat_group , 
                                data=ovarian)
summary(km_fit_trt)

autoplot(km_fit_trt) + 
  ggtitle("Kaplan Meier survival curve based on residual treatment group") +
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```
    
    The survival rate of people in first treatment group is lower than in second group
    
###  ECOG performance status

```{r}
km_fit_ecog <- survfit(Surv(futime, fustat) ~ ecog_status , 
                                data=ovarian)
summary(km_fit_ecog)

autoplot(km_fit_ecog) + 
  ggtitle("Kaplan Meier survival curve based on ECOG performance status") +
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

    The survival rate of people with ECOG performance status-2 is lower than ECOG performance status-1
    
### Divide patients by age, under 60 years old and over 60 years old

    Because mean and median of age is 56.17 and 56.85 years old, the age of separation (60 years) will take a little more than these values
    
```{r}
describe(ovarian$age)
```

```{r}
ovarian_by_age <- mutate(ovarian, age_60 = ifelse((age < 60), 
                                                  "Under60", "Over60"),
              age_60 = factor(age_60))
```

```{r}
km_age_fit <- survfit(Surv(futime, fustat) ~ age_60, data=ovarian_by_age)
autoplot(km_age_fit) + 
  ggtitle("Kaplan Meier survival curve based on age(60 years)") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

    It can be noted that the survival rate of patients under the age of 60 is higher than that of patients over 60.
    
### Create hybrid factor: age60 + treatment group

```{r}
ovarian_by_age$age_treat <- paste(ovarian_by_age$age_60, 
                                           ovarian_by_age$treat_group)
ovarian_by_age$age_treat <- as.factor(ovarian_by_age$age_treat)
```

#### Number of observations in each group

```{r}
ovarian_by_age %>% count(age_treat)
```

```{r}
km_age_treat_fit <- survfit(Surv(futime, fustat) ~ 
                              age_treat, data=ovarian_by_age)
autoplot(km_age_treat_fit) + 
  ggtitle("Kaplan Meier survival curve based on age(60 years) and treatment") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=12, face="bold", hjust = 0.7))
```
    
    The survival rate for people under 60 years is higher in treatment group 2 than in group 1.
    In people over 60 years old, the opposite picture is observed - survival rate in group 1 is higher than in group 2
    
### Create hybrid factor: age60 + ECOG performance status

```{r}
ovarian_by_age$age_ecog <- paste(ovarian_by_age$age_60, 
                                           ovarian_by_age$ecog_status)
ovarian_by_age$age_ecog <- factor(ovarian_by_age$age_ecog)
```

#### Number of observations in each group

```{r}
ovarian_by_age %>% count(age_ecog)
```

```{r}
km_age_ecog_fit <- survfit(Surv(futime, fustat) ~ 
                              age_ecog, data=ovarian_by_age)
autoplot(km_age_ecog_fit) + 
  ggtitle("Kaplan Meier survival curve based on age(60 years) and ECOG performance status") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=10, face="bold", hjust = 0.7))
```

    In all groups of age (under and over 60) survival rate with ECOG performance status=1 higher than ECOG performance status=2

### Create hybrid factor: age60 + residual disease present

```{r}
ovarian_by_age$age_resid <- paste(ovarian_by_age$age_60, 
                                           ovarian_by_age$resid_disease)
ovarian_by_age$age_resid <- factor(ovarian_by_age$age_resid)
```

#### Number of observations in each group

```{r}
ovarian_by_age %>% count(age_resid)
```

    Autoplot can't plot survival plots in case with 1 number of observations in group, let's plot it with ggsurvplot

```{r}
km_age_resid_fit <- survfit(Surv(futime, fustat) ~ 
                              age_resid, data=ovarian_by_age)

ggsurvplot(km_age_resid_fit, data = ovarian_by_age, xlab = "Time in days",
   ylab = "Survival", conf.int = TRUE, ggtheme = theme_bw(), legend.labs =
    c("Over60 No", "Over60 Yes", "Under60 No", "Under60 Yes"),legend = "right",
   linetype = 1, 
   title = "Kaplan Meier survival curve based on age(60 years) 
   and residual disease present"
) 
```

    The survival rate for people under 60 years with residual disease is lower than for people without residual disease
    In people over 60 years old survival rate with residual disease is higher than for people without residual disease
    
    
# Task 2. Which groups differ in terms of survival

## Between residual disease present

```{r}
survdiff(Surv(futime, fustat) ~ resid_disease, data = ovarian)
logrank_test(Surv(futime, fustat) ~ resid_disease, data = ovarian)
```

    Conclusion: survival between groups: residual disease present does not differ:
        - survdiff: Chisq= 3.6  on 1 degrees of freedom, p-value = 0.06 
        - logrank_test: Z = 1.9234, p-value = 0.05444
    
    
## Between treatment group

```{r}
survdiff(Surv(futime, fustat) ~ treat_group, data = ovarian)
logrank_test(Surv(futime, fustat) ~ treat_group, data = ovarian)
```


    Conclusion: survival between treatment group does not differ
        - survdiff: Chisq= 1.1  on 1 degrees of freedom, p-value = 0.3 
        - logrank_test: Z = -1.0296, p-value = 0.3032
    
## Between ECOG performance status

```{r}
survdiff(Surv(futime, fustat) ~ ecog_status, data = ovarian)
logrank_test(Surv(futime, fustat) ~ ecog_status, data = ovarian)
```

    Conclusion: survival between group ECOG performance status does not differ:
        - survdiff: Chisq= 0.5, 1 degrees of freedom, p-value = 0.5
        - logrank_test: Z = 0.69106, p-value = 0.4895
    
    
# Task 3. Assess the factors that influence the risk and assess the risk ratio 

## Analysis of factors affecting survival (Cox model).

```{r, message= FALSE}
cox <- coxph(Surv(futime, fustat) ~ age + resid_disease + treat_group +
               ecog_status, data = ovarian)
summary(cox)

cox_fit <- survfit(cox)

autoplot(cox_fit) + 
  ggtitle("Influence of all factors to survival (Cox model)") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

    Conclusion: of all predictors, only the variable age significantly affects to survival (p-value = 0.00777)
    
## Carry out a covariance analysis to study possible changes in the parameters and their influence over time

```{r, message=FALSE}
cov_fit <- aareg(Surv(futime, fustat) ~ age + resid_disease + treat_group +
               ecog_status, data = ovarian)
autoplot(cov_fit)
```

    It is noticeable that with increasing age, the present of residual disease increases

## Hazard Ratio 

```{r}
fit_coxph <- coxph(Surv(futime, fustat) ~ age + resid_disease + treat_group +
               ecog_status, data = ovarian)
ggforest(fit_coxph, data = ovarian)
```

    Conclusion: as in Cox model of all predictors, only the variable age significantly affects to survival (p-value = 0.008)
    
## Random forest

```{r}
# ranger model
random_fit <- ranger(Surv(futime, fustat) ~ age + resid_disease + treat_group +
               ecog_status, data = ovarian,
                     mtry = 4,
                     importance = "permutation",
                     splitrule = "extratrees",
                     verbose = TRUE)

# Average the survival models
death_times <- random_fit$unique.death.times 
surv_prob <- data.frame(random_fit$survival)
avg_prob <- sapply(surv_prob,mean)

# Plot the survival models for each patient
plot(random_fit$unique.death.times,random_fit$survival[1,], 
     type = "l", 
     ylim = c(0,1),
     col = "red",
     xlab = "Time (days)",
     ylab = "survival",
     main = "Patient Survival Curves")

cols <- colors()
for (n in sample(c(2:dim(ovarian)[1]), 20)){
  lines(random_fit$unique.death.times, random_fit$survival[n,], type = "l", col = cols[n])
}
lines(death_times, avg_prob, lwd = 2)
legend("bottomleft", legend = c('Average = black'))
```

## Comparison of survival charts of all methods

```{r}
km <- rep("Kaplan Meier",length(km_fit$time))
km_df <- data.frame(km_fit$time,km_fit$surv,km)
names(km_df) <- c("Time","Surv","Model")

cox <- rep("Cox",length(cox_fit$time))
cox_df <- data.frame(cox_fit$time,cox_fit$surv,cox)
names(cox_df) <- c("Time","Surv","Model")

rf <- rep("Random forest",length(random_fit$unique.death.times))
rf_df <- data.frame(random_fit$unique.death.times,avg_prob,rf)
names(rf_df) <- c("Time","Surv","Model")

plot_df <- rbind(km_df,cox_df,rf_df)

p <- ggplot(plot_df, aes(x = Time, y = Surv, color = Model))
p + geom_line() + 
  ggtitle("Comparison of survival charts of all models") + 
  xlab("Time (day)") + ylab("Survival") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.7))
```

    Conclusion: all models present similar survival results
    